---
title: "HW3"
author: "Nahomi Farington"
date: "2024-02-13"
output: html_document
---
Group: Amirah, Catherine, Nahomi
## PP1
```{r}
how_many_rolls_pp1 <- 2
how_many_sims_pp1 <- 1

sim_rolls_mat_pp1 <- matrix(sample(1:6, how_many_rolls_pp1 * how_many_sims_pp1, replace = TRUE), nrow = how_many_rolls_pp1)
```

```{r}
if_come_up_5_PP1 <- as.numeric(sim_rolls_mat_pp1 == 5)
mean(if_come_up_5_PP1)
```
## [1] 0.5

```{r}
observed_countspp1 <- table(sim_rolls_mat_pp1 == 5)
print(observed_countspp1)
```
## FALSE  TRUE 
##     1     1

In this protocol, a pair of dice rolls determine unfairness if a 5 comes up, leading to a probability of false positives and negatives, as any instance of a 5 would prompt the conclusion of unfairness.


## PP2
```{r}
how_many_rolls_pp2 <- 2
how_many_sims_pp2 <- 20

sim_rolls_mat_pp2 <- matrix(sample(1:6, how_many_rolls_pp2 * how_many_sims_pp2, replace = TRUE), nrow = how_many_rolls_pp2)
```

```{r}
if_come_up_5_pp2 <- as.numeric(sim_rolls_mat_pp2 == 5)
mean(if_come_up_5_pp2)
```
## [1] 0.125
```{r}
observed_countspp2 <- table(sim_rolls_mat_pp2 == 5)
print(observed_countspp2)
```
## FALSE  TRUE 
##    35     5

In PP2 the dice were rolled 20 times. Unfairness is concluded if a 5 appears more than half of the time. The calculated mean proportion of 0.075 represents the observed frequency of obtaining a 5.  There is a moderate risk of both false positives and false negatives. The observed count shows the dice are fair.


## PP3

```{r}
how_many_rolls_pp3 <- 2
how_many_sims_pp3 <- 100

sim_rolls_mat_pp3 <- matrix(sample(1:6, how_many_rolls_pp3 * how_many_sims_pp3, replace = TRUE), nrow = how_many_rolls_pp3)
```

```{r}
if_come_up_5_pp3 <- as.numeric(sim_rolls_mat_pp3 == 5)
mean(if_come_up_5_pp3)
```
## [1] 0.19

```{r}
observed_countspp3 <- table(sim_rolls_mat_pp3 == 5)
print(observed_countspp3)
```
## FALSE  TRUE 
##   162    38

100 rolls in the third protocol, the criteria for unfairness remain the same: a 5 should appear more than half of the time. The mean proportion of 0.195 signifies the observed frequency of 5s. There is less risk of false positives compared to PP2, there is still a potential for false negatives. The observed counts show the dice are fair.

## EP1
```{r}
how_many_rolls_EP1 <- 20
how_many_sims_EP1 <- 1000

sim_rolls_mat_EP1 <- matrix(sample(1:6, how_many_rolls_EP1 * how_many_sims_EP1, replace = TRUE), nrow = how_many_rolls_EP1)
```

```{r}
if_come_up_5_EP1 <- as.numeric(sim_rolls_mat_EP1 == 5)
mean(if_come_up_5_EP1)
```
## [1] 0.16865

```{r}
observed_countsEP1 <- table(sim_rolls_mat_EP1 == 5)
print(observed_countsEP1)
```
## FALSE  TRUE 
## 16627  3373

The experimental protocol involves 20 rolls in each of 1000 simulations. The criteria for unfairness is the same considering a 5 more than half of the time. The mean proportion of 0.1668 indicates the average frequency of 5s across all simulations. This protocol mitigates the risk of false positives compared to PP2 and pp3 due to the increased number of simulations. Once again, the observed count shows the dice is fair.
