---
title: "Farington_HW7lab6"
author: "Nahomi Farington"
date: "2024-04-03"
output: html_document
---
```{r}
load("C:/Users/Owner/Desktop/Psychology/R-work/V0500_Lecture1/BRFSS2022/.RData")
```

```{r}
library(ggplot2)
library(stargazer)
library(tidyverse)
```


```{r}

brfss22$Age_midpt <- fct_recode(brfss22$X_AGEG5YR, "21" = "Age 18 to 24",
                                      "27" = "Age 25 to 29", "32" = "Age 30 to 34",
                                      "37" = "Age 35 to 39", "42" = "Age 40 to 44",
                                      "47" = "Age 45 to 49", "52" = "Age 50 to 54",
                                      "57" = "Age 55 to 59", "62" = "Age 60 to 64",
                                      "67" = "Age 65 to 69", "72" = "Age 70 to 74",
                                      "77" = "Age 75 to 79", "82" = "Age 80 or older",
                                      NULL = "Dont know/Refused/Missing")
brfss22$Age_midpt <- as.numeric(levels(brfss22$Age_midpt))[brfss22$Age_midpt]
brfss22$income_midpoint <- fct_recode(brfss22$INCOME3, 
                        "7500" = "Household income less than $10,000",
                        "12500" = "Less than $15,000 ($10,000 to less than $15,000)",
                        "17500" = "Less than $20,000 ($15,000 to less than $20,000) ",
                        "22500" = "Less than $25,000 ($20,000 to less than $25,000) ",
                        "30000" = "Less than $35,000 ($25,000 to less than $35,000) ",
                        "42500" = "Less than $50,000 ($35,000 to less than $50,000) ",
                        "62500" = "Less than $75,000 ($50,000 to less than $75,000)",
                        "87500" = "Less than $100,000 ($75,000 to less than $100,000)",
                        "125000" = "Less than $150,000 ($100,000 to less than $150,000)",
                        "175000" = "Less than $200,000 ($150,000 to less than $200,000)",
                        "210000" = "$200,000 or more",
                        NULL = "Dont know/Not sure",
                        NULL = "Refused")
brfss22$income_midpoint <- as.numeric(levels(brfss22$income_midpoint))[brfss22$income_midpoint]

brfss22$Educ_number <- fct_recode(brfss22$EDUCA, 
                                  "0" = "Never attended school or only kindergarten", 
                                  "4.5" = "Grades 1 through 8 (Elementary)",
                                  "10" = "Grades 9 through 11 (Some high school)",
                                  "12" = "Grade 12 or GED (High school graduate)",
                                  "14" = "College 1 year to 3 years (Some college or technical school)",
                                  "16" = "College 4 years or more (College graduate)",
                                  NULL = "Refused" )
brfss22$Educ_number <- as.numeric(levels(brfss22$Educ_number))[brfss22$Educ_number]


brfss_marijan <- "dont know not sure"

select1 <- !is.na(brfss22$MARIJAN1)
brfss_marijan <- subset(brfss22, select1 )


p_cannabis_age <- ggplot(data = brfss_marijan,
                       mapping = aes(x = Age_midpt,
                                     y = MARIJAN1))
p_cannabis_age + geom_smooth()
```

```{r}
model_1 <- lm(MARIJAN1 ~ Age_midpt, data = brfss_marijan)
summary(model_1)
```

```{r}
model_2 <- lm(MARIJAN1 ~ Age_midpt + X_PRACE2 + X_HISPANC + EDUCA, data = brfss_marijan)
summary(model_2)
require(stargazer)
stargazer(model_2, type = "??")
```

#lab 6

```{r}
model_2 <- lm(MARIJAN1 ~ Age_midpt + MARITAL + MSCODE + EDUCA , data = brfss_marijan)
summary(model_2)

model_3 <- lm(MARIJAN1 ~ Age_midpt + I(Age_midpt^2) + MARITAL + MSCODE + EDUCA, data = brfss_marijan)
summary(model_3)
```

```{r}
to_be_predicted_2 <- data.frame(Age_midpt = 25:55, MARITAL = "Married",
                          MSCODE = "in central city", EDUCA = "College 4 years or more (College graduate)")

to_be_predicted_2$yhat <- predict(model_2, newdata = to_be_predicted_2)

summary(to_be_predicted_2$yhat)
```

```{r}
summary
model_4 <- lm(MARIJAN1 ~ Age_midpt*(MARITAL + EDUCA) + I(Age_midpt^2)*(MARITAL + EDUCA) + MARITAL + MSCODE + EDUCA, data = brfss_marijan)
summary(model_4)
```
```{r}
anova(model_2, model_3, model_4)
```


```{r}
to_be_predicted_2$yhat3 <- predict(model_3, newdata = to_be_predicted_2)
to_be_predicted_2$yhat4 <- predict(model_4, newdata = to_be_predicted_2)


d_for_graphing <- data.frame(Age_midpt = 25:55, 
                             to_be_predicted_2$yhat,
                             to_be_predicted_2$yhat3,
                             to_be_predicted_2$yhat4)

p_predvals <- ggplot(d_for_graphing, aes(Age_midpt))
p_predvals + geom_line(aes(y = to_be_predicted_2.yhat)) + 
  geom_line(aes(y = to_be_predicted_2.yhat3), color = 'blue') +
  geom_line(aes(y = to_be_predicted_2.yhat4), color = 'red')
```

```{r}
quantile(brfss_marijan$MARIJAN1, probs = c(0.99,0.95,0.94,0.93,0.92,0.91,0.9))

brfss_marijan$MARIJAN_factor <- as.factor(brfss_marijan$MARIJAN1)
summary(brfss_marijan$MARIJAN_factor)
```

```{r}
brfss_marijan$Marijan_01 <- as.numeric(brfss_marijan$MARIJAN1 > 0)

model_5 <- lm(Marijan_01 ~ Age_midpt*(MARITAL + EDUCA) + I(Age_midpt^2)*(MARITAL + EDUCA) + MARITAL + MSCODE + EDUCA, data = brfss_marijan)
summary(model_5)
```

```{r}
anova(model_4,model_5)
```

1-Markowitz, A. J. (2017). Associations Between School Connection and Depressive Symptoms From Adolescence Through Early Adulthood: Moderation by Early Adversity. Journal of Research on Adolescence (Wiley-Blackwell), 27(2), 298–311. https://doi-org.ccny-proxy1.libr.ccny.cuny.edu/10.1111/jora.12275

The first paper discusses the relationship between school connection and depressive symptoms in youth, particularly those at higher risk for depressive symptoms. The study uses data from the National Longitudinal Study of Adolescent Health and evaluates statistics such as regression coefficients, standard errors, and p-values to analyze associations between school connection, adversity, and depressive symptoms across genders in adolescence.

2-Gunnarsdóttir, H., Hensing, G., & Hammarström, A. (2021). Poor school connectedness in adolescence and adulthood depressiveness: a longitudinal theory-driven study from the Northern Sweden Cohort. European Journal of Public Health, 31(4), 797–802. https://doi-org.ccny-proxy1.libr.ccny.cuny.edu/10.1093/eurpub/ckab027

The paper explores the link between adolescents' school connectedness, family climate, and depressiveness in adulthood using Bronfenbrenner’s ecological theory. It emphasizes the importance of early life experiences in shaping mental health outcomes. The study utilizes data from the Northern Swedish Cohort, involving 481 women and 526 men born in 1965, with data collected at ages 16, 21, 30, and 43. Statistical methods such as generalized linear models with random intercepts were employed to analyze the associations.